# Prerequisites

This repository contains an R package template supporting the use of the
`Rcpp` and `RcppArmadillo` packages. To build the package using the commands
below requires the `devtools` and `roxygen2` packages as well. To find
out if any of these packages are missing from your R installation run

```{r eval=FALSE}
requiredPackages = c("Rcpp", "RcppArmadillo", "devtools", "roxygen2")
missingPackages = setdiff(requiredPackages, rownames(installed.packages()))
```

and to install any missing packages run 

```{r eval=FALSE}
if(length(missingPackages)>0) install.packages(missingPackages)
```

 The package template includes both `R` and
`C++` code, so a
[suitable](https://cran.r-project.org/doc/manuals/r-release/R-admin.html) `C++`
compiler must be available to build the package from source.


### Building and checking the package from within R
```{r eval=FALSE}
Rcpp::compileAttributes()
roxygen2::roxygenize()
# devtools::document() a slower, but more thorough, way to create documentation.
```

### Checking and building and the package from within R
```{r eval=FALSE}
devtools::check(check_dir="build")
# devtools::test() tests are run during check, and can also be run separately.
devtools::build(path="build")
```

### Installing, running, and uninstalling the package from within R
```{r eval=FALE}
devtools::install(upgrade_dependencies=FALSE)
RPackageTemplate::r_hello_world()
devtools::uninstall()
```

### Uninstalling the package from within R
```{r eval=FALSE}