# Prerequisites

This repository contains an R package template supporting use of the
`Rcpp` and `RcppArmadillo` packages. To build the package using the commands
below requires the `devtools`, `roxygen2` and `testthat` packages as well. To
find out if any of these packages are missing from your R installation run:

```{r eval=FALSE}
requiredPackages = c("Rcpp", "RcppArmadillo", "devtools", "roxygen2", "testthat")
missingPackages = setdiff(requiredPackages, rownames(installed.packages()))
```

To install any missing packages run:

```{r eval=FALSE}
if(length(missingPackages)>0) install.packages(missingPackages)
```

The package template includes both `R` and `C++` code, so a
[suitable](https://cran.r-project.org/doc/manuals/r-release/R-admin.html) `C++`
compiler must be available to build the package from source. Windows users can
try the [Rtools](https://cran.r-project.org/bin/windows/Rtools/) software 
collection. Details regarding how to install and configure these tools can be
found on the RStudio website under [Package Development
Prerequisites](https://support.rstudio.com/hc/en-us/articles/200486498-Package-Development-Prerequisites).

Mac users should first install
[xcode](http://itunes.apple.com/us/app/xcode/id497799835?mt=12). Using the
`RcppArmadillo` package also requires a Fortran compiler, which must be
installed separately from xcode. Note the version currently hosted on the R 
project website is outdated and will not work. As described 
[here](http://thecoatlessprofessor.com/programming/rcpp-rcpparmadillo-and-os-x-mavericks-lgfortran-and-lquadmath-error/),
a newer version can be downloaded and installed by opening a terminal and 
executing:

> curl -O http://r.research.att.com/libs/gfortran-4.8.2-darwin13.tar.bz2  
> sudo tar fvxz gfortran-4.8.2-darwin13.tar.bz2 -C /

With the compiler and prerequisite packages installed, the package template can
be prepared, built, and installed by running various R commands. The following
examples assume you have cloned the project template repository to your
computer, and they should be run from within the `RProjectTemplate` directory.

### Preparing the package from within R
```{r eval=FALSE}
Rcpp::compileAttributes()
roxygen2::roxygenize()
# devtools::document() a slower, but more thorough, way to create documentation.
# See http://r-pkgs.had.co.nz/man.html for more about the documentation process.
```

### Checking and building the package from within R
```{r eval=FALSE}
buildDir = "build"
if(!dir.exists(buildDir)) dir.create(buildDir)
devtools::check(check_dir=buildDir)
# devtools::test() tests are run during check, and can also be run separately.
devtools::build(path=buildDir)
```

### Installing and running the package from within R
```{r eval=FALSE}
devtools::install(upgrade_dependencies=FALSE)
RPackageTemplate::r_hello_world()
```

### Uninstalling the package from within R
```{r eval=FALSE}
devtools::uninstall()
```
